Weather covariance matrices as U_hyp for GR50

### 2021-01-14 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Start manuscript writeup with mash env cov matrices

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
library(mashr)
library(rlist)
```

@@

```{r}
mash_plot_covar(m_gulf_gr_no)
mash_plot_covar(m_gulf_gr_hyp)
mash_plot_covar(m_gulf_fl_no)
mash_plot_covar(m_gulf_fl_hyp)

mash_plot_covar(m_midw_gr_no)
mash_plot_covar(m_midw_gr_hyp)
mash_plot_covar(m_midw_fl_no)
mash_plot_covar(m_midw_fl_hyp)

mash_plot_covar(m_gam_gr_no)
mash_plot_covar(m_gam_gr_hyp)
mash_plot_covar(m_gam_fl_no)
mash_plot_covar(m_gam_fl_hyp)
```

@@

cov_gg$ggobject + ggtitle("Gulf greenup 11.4%")
cov_gf$ggobject + ggtitle("Gulf flowering 32.6%")
cov_mg$ggobject + ggtitle("Midwest greenup 35.6%")
cov_mf$ggobject + ggtitle("Midwest flowering 45.7%")
cov_gamg$ggobject + ggtitle("Gulf and Midwest greenup 9.6%")
cov_gamf$ggobject + ggtitle("Gulf and Midwest flowering 46%")
```

46% of effect mass for both subpop flowering was on hyp matrices, while only 9.6% of greenup was.
Gulf greenup: 11.4%  flowering: 32.6%
Midwest greenup: 35.6% flowering 45.7%

```{r}
cov_gamf$covar_df %>% 
  mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~ "data-driven",
                          grepl("single_effect", `Covariance Matrix`) ~ "canonical",
                          grepl("simple_het", `Covariance Matrix`) ~ "canonical",
                          grepl("no_effects", `Covariance Matrix`) ~ "canonical",
                          grepl("identity", `Covariance Matrix`) ~ "canonical",
                          grepl("equal_effects", `Covariance Matrix`) ~ "canonical",
                          grepl("_and_Midwest$", `Covariance Matrix`) ~ "GAM_hyp",
                          grepl("Midwest$", `Covariance Matrix`) ~ "Midw_hyp",
                          grepl("Gulf$", `Covariance Matrix`) ~ "Gulf_hyp",
                          TRUE ~ "other"
                          )) %>%
  group_by(Type) %>%
  summarise(Mass = sum(Mass))

cov_gamg$covar_df %>% 
  mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~ "data-driven",
                          grepl("single_effect", `Covariance Matrix`) ~ "canonical",
                          grepl("simple_het", `Covariance Matrix`) ~ "canonical",
                          grepl("no_effects", `Covariance Matrix`) ~ "canonical",
                          grepl("identity", `Covariance Matrix`) ~ "canonical",
                          grepl("equal_effects", `Covariance Matrix`) ~ "canonical",
                          grepl("_and_Midwest$", `Covariance Matrix`) ~ "hyp",
                          grepl("Midwest$", `Covariance Matrix`) ~ "hyp",
                          grepl("Gulf$", `Covariance Matrix`) ~ "hyp",
                          TRUE ~ "other"
                          )) %>%
  group_by(Type) %>%
  summarise(Mass = sum(Mass))
```
Gulf: 20% canonical, 68.5% data-driven, 9.5% Gulf hyp, 1.6% Midwest hyp
Gulf & Midwest: 28% canonical, 62% data-driven, 5.5% Gulf, 1.2% Midwest, 2.8% G&M

46% of effect mass for both subpop flowering was on hyp matrices, while only 9.6% of greenup was.
Gulf greenup: 11.4%  flowering: 32.6%
Midwest greenup: 35.6% flowering 45.7%

@@

### ii. Gulf
```{r gulf mash results}
length(get_significant_results(gulfm))  # 9142/45931  # 19.9%
get_significant_results(gulfm)[1:10]

covargulf <- mash_plot_covar(m_gulf_fl_hyp, suffix = "19000SNPs_GR50_Gulf", saveoutput = FALSE)
covargulf
get_U_by_mass(gulfm)
mash_plot_Ulist(gulfm, range = get_U_by_mass(gulfm), limits = FALSE)
mash_plot_manhattan_by_condition(gulfm, saveoutput = FALSE)
mash_plot_sig_by_condition(gulfm, saveoutput = FALSE)
```

### ii. G&M
```{r gulf and midwest mash results}
length(get_significant_results(gandm))  # 15104/44765  # 33.7%
get_significant_results(gandm)[1:10]

covargam <- mash_plot_covar(gandm, suffix = "19000SNPs_GR50_Gulf_and_Midwest", saveoutput = TRUE)
covargam
get_U_by_mass(gandm)
mash_plot_Ulist(gandm, range = get_U_by_mass(gandm), limits = FALSE)
mash_plot_manhattan_by_condition(gandm, saveoutput = FALSE)
mash_plot_sig_by_condition(gandm, saveoutput = FALSE)
```

@@

### ii. Plot setup

@@

Get log10BayesFactor for mash SNPs also and combine these into a dataframe for ggplot.
```{r mashhattan plot setup}
m <- gandm
log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
    as.data.frame() %>%
    rownames_to_column(var = "value") %>%
    mutate(value = as.integer(.data$value)) %>%
    as_tibble() %>%
    left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
    dplyr::rename(log10BayesFactor = .data$V1) %>%
    dplyr::select(-.data$value)

#df <- get_topcov(m)

@@

ggman_df <- pw_df %>%
  #left_join(m_meta) %>%
  separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
           extra = "merge", convert = TRUE) %>%
  left_join(log10bf_df, by = "Marker") %>%
  arrange(.data$Chr, .data$Pos)

log10BF <- expression(paste("log"[10], plain("(Bayes Factor)")))

mash_plot_covar(m)
```

Make a list containing all of the ggplots for all entries in the covariance grid represented by `pw_df`.
Then view all the ggplots in gglist.
```{r mashhattan plots loop}
gglist <- list()
for(i in seq_along(names(pw_df)[-(1:2)])){
var <- names(pw_df)[i+2]   
gglist[[i]] <- ggman_df %>%
  arrange(!! sym(var)) %>%
  ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
  switchgrassGWAS::theme_oeco +
  geom_point(aes(color = !! sym(var), fill = !! sym(var),
                 shape = as.factor(.data$Chr))) +
  facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", strip.position = "bottom") +
  scale_color_viridis_c(option = "D", end = 0.95) + 
  scale_fill_viridis_c(option = "D", end = 0.95) +
  theme(strip.text = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 8),
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill=NA)) +
  labs(x = "Chromosome", y = log10BF) +
  scale_x_continuous(expand = c(0.32, 0.32)) +
  scale_shape_manual(values = rep(c(21,22),9), guide = FALSE) +
  geom_hline(yintercept = 2, linetype = 2) +
  ggtitle(var)
}
gglist
```

Which covariance grids have SNPs with log10BF > 2 that load strongly onto them?
For Gulf mash:
GR50_Gulf.6: many, almost all < 2
GR50_Midwest.6: many, almost all < 2
equal_effects.7: many, almost all < 2
GR50_Gulf.6.7: many, almost all < 2
equal_effects.8: some, all <2
ED_PCA_1.9 many, many > 2
ED_PCA_2.9 many, few > 2
ED_PCA_4.9 some, few > 2
ED_tPCA.9: many, many > 2
ED_PCA_3.10: many, few > 2
ED_PCA_5.10: some, few > 2
OK.10 many, some > 2, many >0.8 loading & > 2
SD.10 many, some > 2
cgdd_12c_G&M.10: few, few > 2, some with 80%+ loading here

ED_tPCA.12: many, many > 2, many >0.8 loading esp at high log10BF
SD.12: many, many > 2, many > 0.8 loading esp at high log10BF

TX1.15: few bg, few >2

```{r snps with high sig and high covar loadings}
var <- names(ggman_df)[ncol(ggman_df)-1]  
ggman_df %>%
  pivot_longer(cols = null:var, names_to = "covar_grid", values_to = "loading") %>%
  filter(loading > 0.5) %>%
  group_by(log10BayesFactor > 2, covar_grid, loading > 0.8) %>%
  tally()
```
# -----------
## 3d. Enrichments
•	3d. We looked for enrichments of genes that could potentially affect greenup/flowering in the 20kb genomic windows surrounding the SNPs with significant effects on flowering. 
  •	Also, we looked at other metadata enrichments for these top SNPs.

@@

```{r}
mash_obj <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects33000SNPs_FL50_Midwest_Uhyp.rds")
snp <- snp_attach("~/Github/pvdiv-genome/genetic_subpops/kinship/Pvirgatum_V5_GWAS_363g_Midwest_and_Gulf_subset_maf0.05.rds")

@@

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
library(mashr)
library(rlist)
library(cowplot)
```

@@

@@

cov_gg$ggobject + 
  ggtitle("Gulf greenup", subtitle = "8.4% SNP mass")
cov_gf$ggobject + 
  ggtitle("Gulf flowering", subtitle = "31.8% SNP mass")
cov_mg$ggobject + 
  ggtitle("Midwest greenup", subtitle = "28.6% SNP mass")
cov_mf$ggobject + 
  ggtitle("Midwest flowering", subtitle = "45.1% SNP mass")
cov_gamg$ggobject + 
  ggtitle("Gulf and Midwest greenup", subtitle = "8.5% SNP mass")
cov_gamf$ggobject + 
  ggtitle("Gulf and Midwest flowering", subtitle = "47.3% SNP mass")
```
47.3% of effect mass for both subpop flowering was on hyp matrices, while only 8.5% of greenup was.
Gulf greenup: 8.4%  flowering: 31.8%
Midwest greenup: 28.6% flowering 45.1%

```{r}
get_cov_type <- function(m, summarise = FALSE, subpop = FALSE){
  cov_df <- mash_plot_covar(m)
  if (subpop == FALSE) {
  cov_type <- cov_df$covar_df %>% 
    mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~ "data-driven",
                            grepl("single_effect", `Covariance Matrix`) ~
                              "canonical",
                            grepl("simple_het", `Covariance Matrix`) ~
                              "canonical",
                            grepl("no_effects", `Covariance Matrix`) ~
                              "canonical",
                            grepl("identity", `Covariance Matrix`) ~
                              "canonical",
                            grepl("equal_effects", `Covariance Matrix`) ~
                              "canonical",
                            grepl("_and_Midwest$", `Covariance Matrix`) ~
                              "hypothesized",
                            grepl("Midwest$", `Covariance Matrix`) ~ "hypothesized",
                            grepl("Gulf$", `Covariance Matrix`) ~ "hypothesized",
                            TRUE ~ "other"
                            )) 
  } else {
    cov_type <- cov_df$covar_df %>% 
      mutate(Type = case_when(grepl("ED_", `Covariance Matrix`) ~
                                "data-driven",
                              grepl("single_effect", `Covariance Matrix`) ~
                                "canonical",
                              grepl("simple_het", `Covariance Matrix`) ~
                                "canonical",
                              grepl("no_effects", `Covariance Matrix`) ~
                                "canonical",
                              grepl("identity", `Covariance Matrix`) ~
                                "canonical",
                              grepl("equal_effects", `Covariance Matrix`) ~
                                "canonical",
                              grepl("_and_Midwest$", `Covariance Matrix`) ~
                                "GAM_hyp",
                              grepl("Midwest$", `Covariance Matrix`) ~
                                "Midw_hyp",
                              grepl("Gulf$", `Covariance Matrix`) ~ "Gulf_hyp",
                              TRUE ~ "other"
                              )) 
  }
  if (summarise == TRUE) {
    cov_type <- cov_type %>%
      group_by(Type) %>%
      summarise(Mass = sum(Mass))
  }
  return(cov_type)
}

get_cov_type(m_gulf_gr_hyp, summarise = TRUE, subpop = TRUE)
get_cov_type(m_gulf_fl_hyp, summarise = TRUE, subpop = TRUE)
get_cov_type(m_midw_gr_hyp, summarise = TRUE, subpop = TRUE)
get_cov_type(m_midw_fl_hyp, summarise = TRUE, subpop = TRUE)
get_cov_type(m_gam_gr_hyp, summarise = TRUE, subpop = TRUE)
get_cov_type(m_gam_fl_hyp, summarise = TRUE, subpop = TRUE)
```

```{r make rough draft figure 2}
df_2a <- get_cov_type(m_gulf_gr_hyp, summarise = TRUE) %>%
  mutate(subpop = "Gulf",
         phenotype = "greenup")
df_2b <- get_cov_type(m_midw_gr_hyp, summarise = TRUE) %>%
  mutate(subpop = "Midwest",
         phenotype = "greenup")
df_2c <- get_cov_type(m_gam_gr_hyp, summarise = TRUE) %>%
  mutate(subpop = "Both",
         phenotype = "greenup")
df_2d <- get_cov_type(m_gulf_fl_hyp, summarise = TRUE) %>%
  mutate(subpop = "Gulf",
         phenotype = "flowering")
df_2e <- get_cov_type(m_midw_fl_hyp, summarise = TRUE) %>%
  mutate(subpop = "Midwest",
         phenotype = "flowering")
df_2f <- get_cov_type(m_gam_fl_hyp, summarise = TRUE) %>%
  mutate(subpop = "Both",
         phenotype = "flowering")
df_2gr_b <- df_2a %>% full_join(df_2b) %>% full_join(df_2c)
df_2fl_b <- df_2d %>% full_join(df_2e) %>% full_join(df_2f)

df_2a <- get_cov_type(m_gulf_gr_hyp) %>%
  mutate(subpop = "Gulf",
         phenotype = "greenup")
df_2b <- get_cov_type(m_midw_gr_hyp) %>%
  mutate(subpop = "Midwest",
         phenotype = "greenup")
df_2c <- get_cov_type(m_gam_gr_hyp) %>%
  mutate(subpop = "Both",
         phenotype = "greenup")
df_2d <- get_cov_type(m_gulf_fl_hyp) %>%
  mutate(subpop = "Gulf",
         phenotype = "flowering")
df_2e <- get_cov_type(m_midw_fl_hyp) %>%
  mutate(subpop = "Midwest",
         phenotype = "flowering")
df_2f <- get_cov_type(m_gam_fl_hyp) %>%
  mutate(subpop = "Both",
         phenotype = "flowering")
df_2gr_a <- df_2a %>% full_join(df_2b) %>% full_join(df_2c) %>%
  mutate(`Covariance Matrix` = str_replace(`Covariance Matrix`, "single_effect_Gulf_and_Midwest_", "single_effect_"),
         `Covariance Matrix` = str_replace(`Covariance Matrix`, "single_effect_Midwest_", "single_effect_"),
         `Covariance Matrix` = str_replace(`Covariance Matrix`, "single_effect_Gulf_", "single_effect_")
         )
df_2fl_a <- df_2d %>% full_join(df_2e) %>% full_join(df_2f) %>%
  mutate(`Covariance Matrix` = str_replace(`Covariance Matrix`, "single_effect_Gulf_and_Midwest_", "single_effect_"),
         `Covariance Matrix` = str_replace(`Covariance Matrix`, "single_effect_Midwest_", "single_effect_"),
         `Covariance Matrix` = str_replace(`Covariance Matrix`, "single_effect_Gulf_", "single_effect_")
         )

f2b <- df_2gr_b %>%
  filter(Mass > 1e-05) %>%
  ggplot(aes(x = subpop, y = Mass)) + 
  geom_bar(aes(fill = Type), stat = "identity") +
  scale_fill_viridis_d(end = 0.9) +
  theme(axis.title.x=element_blank(),
        legend.position = "none",
        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45))
f2d <- df_2fl_b %>%
  filter(Mass > 1e-05) %>%
  ggplot(aes(x = subpop, y = Mass)) + 
  geom_bar(aes(fill = Type), stat = "identity") +
  scale_fill_viridis_d(end = 0.9) +
  theme(axis.title.x=element_blank(),
        legend.position = "none",
        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45))
f2a <- df_2gr_a %>%
  filter(Mass > 1e-06) %>%
  ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
  geom_bar(aes(fill = Type), stat = "identity") +
  facet_wrap(~subpop) +
  coord_flip() + scale_fill_viridis_d(end = 0.9) +
  theme(legend.position = c(-0.30, 0.3),
        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
        panel.spacing.x = unit(-0, 'cm'))
f2c <- df_2fl_a %>%
  filter(Mass > 1e-06) %>%
  ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
  geom_bar(aes(fill = Type), stat = "identity") +
  facet_wrap(~subpop) +
  coord_flip() + scale_fill_viridis_d(end = 0.9) +
  theme(legend.position = "none",
        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
        panel.spacing.x = unit(-0, 'cm'))

fig2 <- plot_grid(f2a, f2b, f2c, f2d, nrow = 2, rel_widths = c(0.75, 0.25), rel_heights = c(0.53, 0.47))
save_plot("../manuscript/plots/Figure_2_mash_covariance_masses.png", base_height = 8, base_width = 7, plot = fig2)
```
47.3% of effect mass for both subpop flowering was on hyp matrices, while only 8.5% of greenup was.
Gulf greenup: 8.4%  flowering: 31.8%
Midwest greenup: 28.6% flowering 45.1%

@@

### ii. Gulf
```{r gulf mash results}
length(get_significant_results(m_gulf_gr_hyp))  # 9645/45931  # 20.999%
length(get_significant_results(m_gulf_gr_no))   # 10427/45931 # 22.7%
get_significant_results(m_gulf_gr_hyp)[1:10]

get_U_by_mass(m_gulf_gr_hyp)
mash_plot_Ulist(m_gulf_gr_hyp, range = get_U_by_mass(m_gulf_gr_hyp), limits = FALSE)
mash_plot_manhattan_by_condition(m_gulf_gr_hyp, saveoutput = FALSE)
mash_plot_sig_by_condition(m_gulf_gr_hyp, saveoutput = FALSE)
```

### ii. G&M
```{r gulf and midwest mash results}
length(get_significant_results(m_gam_gr_hyp))  # 15786/44765  # 35.26
length(get_significant_results(m_gam_gr_no))  # 15873/44765  # 35.46%
get_significant_results(m_gam_gr_hyp)[1:10]

get_U_by_mass(m_gam_gr_hyp)
mash_plot_Ulist(m_gam_gr_hyp, range = get_U_by_mass(m_gam_gr_hyp), limits = FALSE)
mash_plot_manhattan_by_condition(m_gam_fl_hyp, saveoutput = FALSE)
mash_plot_sig_by_condition(m_gam_fl_hyp, saveoutput = FALSE)
```

@@

### ii. Functions for posterior weights

@@

Get log10BayesFactor for mash SNPs also and combine these into a dataframe for ggplot.
```{r mashhattan plot setup}
get_post_weights <- function(m){
  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      rownames_to_column(var = "value") %>%
      mutate(value = as.integer(.data$value)) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      dplyr::select(-.data$value)

@@

  ggman_df <- pw_df %>%
    #left_join(m_meta) %>%
    separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
             extra = "merge", convert = TRUE) %>%
    left_join(log10bf_df, by = "Marker") %>%
    arrange(.data$Chr, .data$Pos) 
  return(ggman_df)
}

pw_gg <- get_post_weights(m_gulf_gr_hyp)
pw_gf <- get_post_weights(m_gulf_fl_hyp)

pw_mg <- get_post_weights(m_midw_gr_hyp)
pw_mf <- get_post_weights(m_midw_fl_hyp)

pw_gamg <- get_post_weights(m_gam_gr_hyp)
pw_gamf <- get_post_weights(m_gam_fl_hyp)

get_pw_filtered <- function(m, pw_thresh = 0.5, BF_thresh = 2){
  ggman_df <- get_post_weights(m)
  var <- names(ggman_df)[ncol(ggman_df)-1]  
  pw_summary <- ggman_df %>%
    pivot_longer(cols = null:var, names_to = "covar_grid", 
                 values_to = "loading") %>%
    filter(log10BayesFactor > BF_thresh) %>%
    group_by(covar_grid, loading > pw_thresh) %>%
    tally()
return(pw_summary)
}

get_pw_filtered(m_gulf_gr_hyp, pw_thresh = 0.1, BF_thresh = 2) %>% 
  mutate(Type = case_when(grepl("ED_", covar_grid) ~ "data-driven",
                          grepl("Stand", covar_grid) ~ "canonical",
                          grepl("simple_het", covar_grid) ~ "canonical",
                          grepl("null", covar_grid) ~ "canonical",
                          grepl("identity", covar_grid) ~ "canonical",
                          grepl("equal_effects", covar_grid) ~ "canonical",
                          grepl("_and_Midwest", covar_grid) ~ "hyp",
                          grepl("Midwest", covar_grid) ~ "hyp",
                          grepl("Gulf", covar_grid) ~ "hyp",
                          TRUE ~ "other"
                          ))

get_significant_results(m_gulf_gr_hyp)[1:10]

```
Not many SNPs with U_hyp loading > 10% and SNP with log10BF > 2:
Both GR: 10 & 9 SNPs   cgdd_12c_18d_Gulf_and_Midwest x2 (0.9 and 0.10)
Both FL: 2 SNPs        dyln_change_sec_Midwest

Gulf FL: 0 SNPs
Gulf GR: 11 & 2 SNPs   cgdd_12c_18d_Gulf_and_Midwest.10 & tave_18d_Gulf_and_Midwest.6

Midw GR: 0 SNPs
Midw FL: 11 SNPs       cgdd_12c_gr2fl_Gulf_and_Midwest

### iii. Plot setup
Make a list containing all of the ggplots for all entries in the covariance grid represented by `pw_df`.
Then view all the ggplots in gglist.
```{r mashhattan plots loop}
pw_df <- pw_gg

gglist <- list()
log10BF <- expression(paste("log"[10], plain("(Bayes Factor)")))
for(i in seq_along(names(pw_df)[-(1:5)])){
  var <- names(pw_df)[i+4]  
  gglist[[i]] <- pw_df %>%
    arrange(!! sym(var)) %>%
    ggplot(aes(x = .data$Pos, y = .data$log10BayesFactor)) +
    switchgrassGWAS::theme_oeco +
    geom_point(aes(color = !! sym(var), fill = !! sym(var),
                   shape = as.factor(.data$Chr))) +
    facet_wrap(~ .data$Chr, nrow = 1, scales = "free_x", 
               strip.position = "bottom") +
    scale_color_viridis_c(option = "D", end = 0.95) + 
    scale_fill_viridis_c(option = "D", end = 0.95) +
    theme(strip.text = element_text(size = 8),
          axis.text.y = element_text(size = 8),
          axis.text.x = element_blank(),
          legend.text = element_text(size = 8),
          axis.ticks.x = element_blank(),
          panel.background = element_rect(fill=NA)) +
    labs(x = "Chromosome", y = log10BF) +
    scale_x_continuous(expand = c(0.32, 0.32)) +
    scale_shape_manual(values = rep(c(21,22),9), guide = FALSE) +
    geom_hline(yintercept = 2, linetype = 2) +
    ggtitle(var)
}
gglist
```

# -----------
## 3d. Enrichments
•	3d. We looked for enrichments of genes that could potentially affect greenup/flowering in the 20kb genomic windows surrounding the SNPs with significant effects on flowering. 
  •	Also, we looked at other metadata enrichments for these top SNPs.

@@

### Rice/thaliana FT

anno_df: All are above a 10% FDR. If there are more than 500 SNPs above the FDR, only returns the top 500 by p value. Only gets the closest gene to each SNP. Should be three site sets and three subpop sets, several (four?) flowering phenotypes.

os_annos: OsGene and AtGene columns of this df indicate if the gene has a homolog in rice or A. thaliana with a functional study. Key is the locusName column.
```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
ftanno <- read_csv(file.path(workingdir, "data/Pvirgatum_v5.1.annotation_FTGeneHomologs.csv"))
ftkey <- read_delim(file.path(workingdir, "data", "OsGene_keyword_table.txt"),
                    delim = "\t")
ftanno %>%
  filter(!is.na(OsGene)) %>%
  left_join(ftkey, by = c("OsGene" = "Symbol")) %>%
  filter(Keyword %in% c("flowering time", "heading date")) %>% unique()

osgene_keyword_df <- ftkey %>%
  dplyr::select(-RAPdb, -MSU) %>%
  group_by(Symbol, Keyword) %>%
  dplyr::slice(1) %>%
  pivot_wider(names_from = Keyword, values_from = Title)

osgene_key <- osgene_keyword_df %>%
  dplyr::select(Symbol, `flowering time`, `heading date`, anther, flower,
                `floral meristem`, `floral meristem determinacy`, flowering,
                `flower development`, `carpel differentiation`, stamen, 
                `stamen number`, everything()) %>%
  pivot_longer(cols = -Symbol, names_to = "Keyword", values_to = "Title") %>%
  filter(!is.na(Title)) %>%
  pivot_wider(names_from = Keyword, values_from = Title)

os_annos <- ftanno %>%
  filter(OsGene %in% osgene_key$Symbol | !is.na(AtGene)) %>%
  mutate(Pos_Mb = round(start/10000)/100,
         #Pos_Mb_e = floor(start/10000)/100,
         POS_bin = round(Pos_Mb*100/2)) %>%# ,
         #POS_bin_hi = round(Pos_Mb_e*100/2))
  dplyr::select(OsGene, AtGene, locusName, everything()) %>%
  dplyr::rename(`Gene ID` = locusName)
1599 / 83260
```

# Annotations with rice/thaliana functional annotation papers for specific SNPs
```{r}
library(AnnotationDbi)
txdb <- loadDb("~/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite")
not_all_na <- function(x) any(!is.na(x))
not_any_na <- function(x) all(!is.na(x))
region_size <- 20000

gg_annos <- pw_gg %>% 
  filter((cgdd_12c_18d_Gulf_and_Midwest.10 > .1 | tave_18d_Gulf_and_Midwest.6 > 0.1) & log10BayesFactor > 2) %>%
  mutate(CHR = Chr,
         start = Pos - (region_size/2),
         end = Pos + (region_size/2)) %>%
  pvdiv_table_topsnps(., type = "table", txdb = txdb)

gg_annos %>%
  inner_join(os_annos, by = c("Gene ID")) %>%
  left_join(osgene_key, by = c("OsGene" = "Symbol")) %>%
  dplyr::select(where(not_all_na))

top_annos <- pw_gg %>%
  filter(log10BayesFactor > 2) %>%
  mutate(CHR = Chr,
         start = Pos - (region_size/2),
         end = Pos + (region_size/2)) %>%
  pvdiv_table_topsnps(., type = "table", txdb = txdb) %>%
  inner_join(os_annos, by = c("Gene ID")) %>%
  left_join(osgene_key, by = c("OsGene" = "Symbol")) %>%
  dplyr::select(where(not_all_na)) %>%
  group_by(Chr, POS, `Gene ID`) %>%
  dplyr::slice(1)
```

```{r function}
library(AnnotationDbi)
not_all_na <- function(x) any(!is.na(x))
txdb <- loadDb("~/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite")

get_post_weights <- function(m){
  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      rownames_to_column(var = "value") %>%
      mutate(value = as.integer(.data$value)) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      dplyr::select(-.data$value)

  ggman_df <- pw_df %>%
    #left_join(m_meta) %>%
    separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
             extra = "merge", convert = TRUE) %>%
    left_join(log10bf_df, by = "Marker") %>%
    arrange(.data$Chr, .data$Pos) 
  return(ggman_df)
}

get_post_annos <- function(m, region_size = 20000, BFthresh = 2){
  pw_df <- get_post_weights(m)
  pw_annos <- pw_df %>%
    filter(log10BayesFactor > BFthresh) %>%
    mutate(CHR = Chr,
           start = Pos - (region_size/2),
           end = Pos + (region_size/2)) %>%
    pvdiv_table_topsnps(., type = "table", txdb = txdb) %>%
    inner_join(os_annos, by = c("Gene ID")) %>%
    left_join(osgene_key, by = c("OsGene" = "Symbol")) %>%
    dplyr::select(where(not_all_na)) %>%
    group_by(Chr, POS, `Gene ID`) %>%
    dplyr::slice(1) %>% 
    arrange(desc(log10BayesFactor))
  return(pw_annos)
}
```

```{r}
annos_gg <- get_post_annos(m_gulf_gr_hyp)
annos_gf <- get_post_annos(m_gulf_fl_hyp)
annos_mg <- get_post_annos(m_midw_gr_hyp)
annos_mf <- get_post_annos(m_midw_fl_hyp)
annos_gamg <- get_post_annos(m_gam_gr_hyp)
annos_gamf <- get_post_annos(m_gam_fl_hyp)
```

```{r}
annos_gg
```

### Get mash SNP metadata
```{r}
mash_obj <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects33000SNPs_FL50_Midwest_Uhyp.rds")
snp <- snp_attach("~/Github/pvdiv-genome/genetic_subpops/kinship/Pvirgatum_V5_GWAS_363g_Midwest_and_Gulf_subset_maf0.05.rds")

@@

Example to modify from a previous script:
```{r}
get_U_by_mass(m)
i=2
names(Ulist)[i]
Ulist <- switchgrassGWAS:::get_Ulist(m)
b1 <- svd(Ulist[[i]])
b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 11.5
b1$d[2] / sum(b1$d) # get to 60%
b1$d[3] / sum(b1$d) # 
b1$d[4] / sum(b1$d) # 
b1$d[5] / sum(b1$d) #

sum(b1$d[1:2]) / sum(b1$d) # First 5 eigenvectors explain 1% or more variance, and 27.7% of total variance.
for(g in 1:5){
barplot(b1$v[,g]/Ulist[[1]][which.max(abs(Ulist[[1]][,g])),g])
}

ggb1 <- tibble(`First eigenvector (PVE = 100%)` = b1$v[,1]/Ulist[[1]][which.max(abs(Ulist[[1]][,1])),1],
       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
  separate(Phenotype, into = c("Function", "Garden"), sep = -3) %>%
  mutate(Function = str_remove(Function, "_$"),
         Garden = str_remove(Garden, "^_"))
ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
ggb1_plot <- ggb1 %>%
  ggplot(aes(x = Garden, y = `First eigenvector (PVE = 100%)`)) +
  geom_bar(aes(group = Function, color = Function, fill = Function), stat = "identity", position = "dodge") +
  theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(hjust = 1, angle = 45)) +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_color_viridis_d(option = "B", end = 0.7, 
                      direction = -1) +
  scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, direction = -1) 
save_plot(paste0("../manuscript/plots/SVD_of_greenup_Gulf_", names(Ulist)[i], ".png"), base_height = 4, plot = ggb1_plot)

ggb2 <- tibble(`First eigenvector (PVE = 21.7%)` = b1$v[,2]/Ulist[[1]][which.max(abs(Ulist[[1]][,2])),2],
       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
  separate(Phenotype, into = c("Function", "Garden"), sep = -3) %>%
  mutate(Function = str_remove(Function, "_$"),
         Garden = str_remove(Garden, "^_"))
ggb2$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
ggb2_plot <- ggb2 %>%
  ggplot(aes(x = Garden, y = `First eigenvector (PVE = 21.7%)`)) +
  geom_bar(aes(group = Function, color = Function, fill = Function), stat = "identity", position = "dodge") +
  theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(hjust = 1, angle = 45)) +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_color_viridis_d(option = "B", end = 0.7, 
                      direction = -1) +
  scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, direction = -1) 
save_plot(paste0("../manuscript/plots/SVD_of_greenup_Gulf_", names(Ulist)[i], "Eigenvector_2.png"), base_height = 4, plot = ggb2_plot)

@@

`get_GxE` and `mash_plot_pairwise_sharing()` should do the trick here.
